---
title: "Regression Methods Project"
format: html
editor: visual
---

#### **1. Introduction**

-   **Contextualize the Problem:**

    -   Briefly describe the phenomenon under investigation.

    -   Mention previous findings or theories (e.g., the original Bayesian analysis).

-   **State Objectives:**

    -   Clearly define the aims of your analysis.

-   **Outline:**

    -   Provide a roadmap of your report's structure.

#### **2. Data Description and EDA**

**Data Description**

The dataset contains aggregated results from an experiment involving coin flips conducted by 48 participants using 211 different coins. Each row in the dataset represents the outcomes of flipping a specific coin by a participant under two starting conditions: when the coin starts heads up and when it starts tails up. The variables in the dataset are as follows:

-   **heads_heads**: Number of flips where the coin started heads up and landed heads up.

-   **tails_heads**: Number of flips where the coin started tails up and landed heads up.

-   **N_start_heads_up**: Total number of flips where the coin started heads up.

-   **N_start_tails_up**: Total number of flips where the coin started tails up.

-   **person**: Identifier for the participant.

-   **coin**: Identifier for the coin used.

```{r, message=FALSE, warning=FALSE}
#| echo: false
# Load necessary libraries
library(tidyverse)
library(glue)

# Read the data
df <- read_csv("/Users/graceaverell/Desktop/EPFL/regression methods/project/coin data/analyses/data-agg.csv") %>%
  mutate(
    person = factor(person),
    coin = factor(coin)
  )

```

**Exploratory Data Analysis**

Our analysis begins with an examination of the key probabilities and distributions in the dataset to understand its main features and inform our modeling approach.

**Overall Probability of Landing on the Same Side**

We calculate the overall probability that a coin lands on the same side it started by dividing the total number of same-side outcomes by the total number of flips. The summary statistics are:

```{r, message=FALSE, warning=FALSE}
#| echo: false
# Calculate flips where the coin landed on the opposite side
df <- df %>%
  mutate(
    heads_tails = N_start_heads_up - heads_heads,
    tails_tails = N_start_tails_up - tails_heads
  )

# Calculate total flips and probabilities
df <- df %>%
  mutate(
    total_flips = N_start_heads_up + N_start_tails_up,
    total_same_side = heads_heads + tails_tails,
    prob_same_side = total_same_side / total_flips,
    prob_heads_to_heads = heads_heads / N_start_heads_up,
    prob_tails_to_tails = tails_tails / N_start_tails_up
  )

# Summary statistics for probability of landing on the same side}
summary_stats <- df %>%
  summarise(
    mean_prob_same_side = mean(prob_same_side, na.rm = TRUE),
    median_prob_same_side = median(prob_same_side, na.rm = TRUE),
    sd_prob_same_side = sd(prob_same_side, na.rm = TRUE)
  )

print(summary_stats)
```

The mean probability is **0.508**, slightly above the expected **0.5** for a fair coin, suggesting a minor bias toward landing on the same side. We visualize this distribution in Figure 1.

```{r, message=FALSE, warning=FALSE}
#| echo: false
# Histogram of probabilities of landing on the same side
ggplot(df, aes(x = prob_same_side)) +
  geom_histogram(binwidth = 0.001, fill = "skyblue", color = "black") +
  labs(
    title = "Distribution of Probability of Landing on the Same Side",
    x = "Probability",
    y = "Frequency"
  ) +
  theme_minimal()

```

Figure 1: Histogram of the probability of landing on the same side across all coins and participants.

The histogram shows probabilities clustered around **0.5**, with a slight skew toward higher values, reinforcing the observation of a minor bias.

**Participant-Level Analysis**

To assess the influence of individual flipping techniques, we compute the probability of landing on the same side for each participant:

```{r, message=FALSE, warning=FALSE}
#| echo: false
# Calculate participant-level probabilities
participant_probs <- df %>%
  group_by(person) %>%
  summarise(
    total_heads_heads = sum(heads_heads),
    total_tails_tails = sum(tails_tails),
    total_heads_up = sum(N_start_heads_up),
    total_tails_up = sum(N_start_tails_up),
    total_same_side = total_heads_heads + total_tails_tails,
    total_flips = total_heads_up + total_tails_up,
    prob_same_side = total_same_side / total_flips
  )

# Summary statistics at participant level
participant_summary <- participant_probs %>%
  summarise(
    mean_prob = mean(prob_same_side),
    median_prob = median(prob_same_side),
    sd_prob = sd(prob_same_side)
  )

print(participant_summary)

```

The mean probability at the participant level is **0.510**, slightly higher than the overall mean, indicating some participants are more prone to this bias. The variability among participants is depicted in Figure 2.

```{r, message=FALSE, warning=FALSE}
#| echo: false
# Boxplot of probabilities by participant
ggplot(participant_probs, aes(x = reorder(person, prob_same_side), y = prob_same_side)) +
  geom_boxplot(fill = "lightgreen") +
  labs(
    title = "Probability of Landing on the Same Side by Participant",
    x = "Participant",
    y = "Probability"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Figure 2: Probability of landing on the same side by participant.

The boxplot reveals that while most participants have probabilities near **0.5**, some exhibit higher values, suggesting individual differences may influence outcomes. We also see some obvious outliers.

**Coin-Level Analysis**

We analyze probabilities at the coin level to determine if certain coins are more likely to land on the same side:

```{r, message=FALSE, warning=FALSE}
#| echo: false
# Calculate coin-level probabilities
coin_probs <- df %>%
  group_by(coin) %>%
  summarise(
    total_heads_heads = sum(heads_heads),
    total_tails_tails = sum(tails_tails),
    total_heads_up = sum(N_start_heads_up),
    total_tails_up = sum(N_start_tails_up),
    total_same_side = total_heads_heads + total_tails_tails,
    total_flips = total_heads_up + total_tails_up,
    prob_same_side = total_same_side / total_flips
  )

# Summary statistics at coin level
coin_summary <- coin_probs %>%
  summarise(
    mean_prob = mean(prob_same_side),
    median_prob = median(prob_same_side),
    sd_prob = sd(prob_same_side)
  )

print(coin_summary)

```

The mean probability is **0.504**, slightly lower than the overall mean. Figure 3 illustrates the distribution of probabilities across coins.

```{r, message=FALSE, warning=FALSE}
#| echo: false
# Histogram of probabilities by coin
ggplot(coin_probs, aes(x = prob_same_side)) +
  geom_histogram(binwidth = 0.001, fill = "salmon", color = "black") +
  labs(
    title = "Distribution of Probability of Landing on the Same Side by Coin",
    x = "Probability",
    y = "Frequency"
  ) +
  theme_minimal()

```

The histogram shows a tighter clustering around **0.5**, indicating less variability among coins compared to participants, and suggesting coin characteristics have minimal impact.

**Effect of Starting Side**

We investigate whether the starting side affects the probability of landing on the same side:

```{r, message=FALSE, warning=FALSE}
#| echo: false
# Compare probabilities based on starting side
starting_side_probs <- df %>%
  summarise(
    mean_prob_heads_to_heads = mean(prob_heads_to_heads, na.rm = TRUE),
    mean_prob_tails_to_tails = mean(prob_tails_to_tails, na.rm = TRUE),
    sd_prob_heads_to_heads = sd(prob_heads_to_heads, na.rm = TRUE),
    sd_prob_tails_to_tails = sd(prob_tails_to_tails, na.rm = TRUE)
  )

print(starting_side_probs)

```

Both starting sides have a mean probability of **0.508**, indicating a consistent slight bias regardless of the starting orientation. Figure 4 compares the distributions.

```{r, message=FALSE, warning=FALSE}
#| echo: false
# Prepare data for plotting
df_long <- df %>%
  select(prob_heads_to_heads, prob_tails_to_tails) %>%
  pivot_longer(
    cols = c(prob_heads_to_heads, prob_tails_to_tails),
    names_to = "starting_side",
    values_to = "probability"
  ) %>%
  mutate(
    starting_side = recode(starting_side,
                           prob_heads_to_heads = "Heads Up",
                           prob_tails_to_tails = "Tails Up")
  )

# Boxplot of probabilities by starting side
ggplot(df_long, aes(x = starting_side, y = probability, fill = starting_side)) +
  geom_boxplot() +
  labs(
    title = "Probability of Landing on the Same Side by Starting Side",
    x = "Starting Side",
    y = "Probability"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

Figure 4: Probability of landing on the same side by starting side.

The boxplot shows similar distributions for both starting sides, suggesting the starting orientation has a negligible effect.

**Discussion of EDA Findings**

Our exploratory analysis reveals:

1.  **Slight Overall Bias:** A mean probability of **0.508** indicates a minor bias toward coins landing on the same side they started.

2.  **Participant Variability:** The higher mean probability (**0.510**) and greater variability among participants suggest individual flipping techniques may influence outcomes.

3.  **Minimal Coin Influence:** The mean probability at the coin level (**0.504**) is close to **0.5**, with less variability, indicating coin characteristics have little effect.

4.  **Negligible Starting Side Effect:** Similar probabilities for both starting sides imply that the starting orientation does not significantly impact the outcome.

**Implications for Modeling**

-   **Include Participant Effects:** Given the variability among participants, incorporating participant-specific random effects in our models is warranted.

-   **Consider Starting Side:** Although the effect is minimal, including starting side as a predictor may capture subtle influences.

-   **Coin Effects May Be Limited:** The minimal variability at the coin level suggests coin-specific effects may not be critical but could be included.

```{r, message=FALSE, warning=FALSE}
#| echo: false

```

Example structure for the rest:

#### **3. Methodology**

-   **Justify Model Choices:**

    -   Explain why certain models are appropriate (e.g., mixed-effects models for nested data).

-   **Detail Statistical Methods:**

    -   Clearly present the statistical models and assumptions.

-   **Reference Statistical Theory:**

    -   Cite relevant textbooks or papers for the methods used.

#### **4. Results and Discussion**

-   **Present Model Results:**

    -   Tables of coefficients, standard errors, p-values.

-   **Interpret Findings:**

    -   Discuss what the results mean in the context of the research question.

-   **Include Visualizations:**

    -   Predicted probabilities, effect plots, residual diagnostics.

-   **Discuss Limitations:**

    -   Address any potential issues with the analysis.

#### **5. Conclusions**

-   **Summarize Key Findings:**

    -   Recap the main results and their implications.

-   **Future Directions:**

    -   Suggest how the analysis could be extended or improved.

```{r, message=FALSE, warning=FALSE}
#| echo: false

```

The `echo: false` option disables the printing of code (only output is displayed).
